---
title: "Using SDMtune for MaxEnt with Cross-Validation"
subtitle: "Plot 1"
author: "Briana Barajas"
date: 2024-03-22
---

```{r}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r packages, include=FALSE}
# wrangling
library(tidyverse)
library(here)
library(janitor)

# raster/geospatial
library(sf)
library(tmap)
library(terra)

# modeling
library(SDMtune)
library(rsample)
library(rJava)

# set data directory
output_dir <- here("data", "updated-rasters")
data_dir <- here("data", "CampRoberts_spatial_data")
```


## Read in Data

```{r, results='hide'}
# read environmental rasters
canopy <- rast(here(output_dir, "canopy", "plot1", "canopy.asc"))
dnd_dn <- rast(here(output_dir, "dnd_dn", "plot1", "dnd_dn.asc"))
elev <- rast(here(output_dir, "elev", "plot1", "elev.asc"))
gs_dn <- rast(here(output_dir, "gs_dn", "plot1", "gs_dn.asc"))
hli <- rast(here(output_dir, "hli", "plot1", "hli.asc"))
li_dn <- rast(here(output_dir, "li_dn", "plot1", "li_dn.asc"))
slope <- rast(here(output_dir, "slope", "plot1", "slope.asc"))


# combine environmental rasters
predictor_stack_rast <- c(elev, hli, slope, canopy, dnd_dn, li_dn, gs_dn)

# read in occurence data
occurences <- st_read(here(data_dir, "Species_pts", "CR_BASP_obs_11Jul22.shp")) %>% 
  st_make_valid() %>% 
  clean_names()
```


## Maxent Using AUC
For this run, I tuned all the hyper parameters for Maxent as it is done in the `{SDMtune}` documentation. I added cross validation folds, to the standards `gridSearch` function and compared it to the `optimizeModel` function.

```{r}
set.seed(479)

## ========================================
##             Prepare Data            ----
## ========================================
# update df, isolating occurence lat and long
occurence_coords <- occurences %>% 
  st_drop_geometry() %>% 
  rename(y = latitude,
         x = longitude) %>% 
  dplyr::select(x,y)

# create background points using raster stack
bg_points <- spatSample(predictor_stack_rast,
                        size = 1000,
                        replace = TRUE,
                        xy = TRUE)

# isolate coordinates of bg points
bg_coords <- bg_points %>% dplyr::select(c(x,y))

## ========================================
##            Maxnet Pre-processing    ----
## ========================================

# create SWD object using data
swd_obj <- prepareSWD(species = "CA Slender Salamander",
                      p = occurence_coords, # occurence points
                      a = bg_coords, # background point coordinates
                      env = predictor_stack_rast)

# split data into test and train
split <- trainValTest(swd_obj, 
                      test = 0.2, # proportion of data for testing
                      only_presence = FALSE, #includes bg pts in each partition
                      seed = 4) # keep same random split
train <- split[[1]]
test <- split[[2]]

# prepare cross validation folds
cv_folds <- randomFolds(train, k = 4, only_presence = TRUE)


## ========================================
##           Maxnet Model Tuning       ----
## ========================================
# select hyper parameters for testing
param_tune <- list(
  reg = seq(0.1, 3, 0.1), # regularization multiplier
  fc = c("lq", "lh", "lqp", "lqph", "lqpht")) # feature class combination

# train maxnet model
maxnet_model <- train(method = "Maxnet",
                      folds = cv_folds,
                      data = train)

# # remove variebles with importance less than 2% IF it doesn't decrease model performance
# # AUC improves by only 0.01 when using reduced variable mode
# maxnet_model_red <- reduceVar(maxnet_model,
#                     th = 2,
#                     metric = "auc",
#                     test = test,
#                     use_jk = TRUE)

## ========================================
##           Maxnet Model Results      ----
## ========================================
# try combinations with gridSearch
grid_res <- gridSearch(maxnet_model,
                       hypers = param_tune,
                       metric = "auc",
                       test = test)

# view results
head(grid_res@results[order(-grid_res@results$test_AUC), ])



```

## Maxent Using AICC
For this model, I will be using "AICc" instead of "AUC". Note that AICc can only be used for Maxent models, and is not indicative of predictive performance. However, given the small sample size in many of these plots 

```{r}
# train maxnet model
maxnet_model2 <- train(method = "Maxnet",
                      data = train)

# find best hyper parameter combinations
grid_res2 <- gridSearch(maxnet_model2,
                       hypers = param_tune,
                       metric = "aicc",
                       env = predictor_stack_rast,
                       save_models = FALSE)

# view results
head(grid_res2@results[order(grid_res2@results$AICc), ])

View(grid_res2@results)

```

## Running the Boosted Regression Model

```{r}
## ========================================
##           BRT Model Tuning       ----
## ========================================
# select hyper parameters for testing
brt_param_tune <- list(
  n.trees = seq(10, 300, 10))

# train maxnet model
brt_model <- train(method = "BRT",
                      folds = cv_folds,
                      data = train)

## ========================================
##           BRT Model Results      ----
## ========================================
# try combinations with gridSearch
brt_grid_res <- gridSearch(brt_model,
                       hypers = brt_param_tune,
                       metric = "auc",
                       test = test)

# view results
head(brt_grid_res@results[order(-brt_grid_res@results$test_AUC), ])

head(grid_res@results[order(-grid_res@results$test_AUC), ])


```





