---
title: "Boosted Regression Tree"
author: "Briana Barajas"
date: 2024-06-03
---

## Preparation

```{r}
rm(list = ls())

# wrangling
library(tidyverse)
library(here)
library(janitor)

# raster/geospatial
library(sf)
library(tmap)
library(terra)

# modeling
library(SDMtune)
library(rsample)
library(rJava)

# source functions
source(here("R", "sdmtune-brt-function.R"))

```

## Plot Summary

Each plot indicates a different transect where cover board surveys were conducted, and contains it's own unique values of occurrence points. The maximum entropy model is often noted for it's capabilities even when given a small data set. Since each plot has a unique number of occurrence points, some of the analyses conducted compare plots with low vs. high numbers of points to see how the AUC was affected.

| Plot Number | Number of Occurrences |
|-------------|----------------------|
| 1           | 51                   |
| 2           | 57                   |
| 3           | 78\*                 |
| 4           | 19                   |
| 5           | 72                   |
| 6           | 39                   |
| 7           | 10\*                 |
| 8           | 21                   |


## All Variables - Iterate Through Plots 1-8

```{r, eval=FALSE}

# set data directory
rast_dir <- here("data", "raster-stacks", "full-stack")
point_dir <- here("data", "CampRoberts_spatial_data")

# create empty df
brt_all_predictions <- data.frame()


# run BRT for all 8 plots (all & reduced variables)
for (i in seq(1:2)) {
  
  # .............run both BRT models.............
  brt_res <- tune_brt(plot_number = i,
                      point_dir = point_dir,
                      rast_dir = rast_dir,
                      include_variables = "both")
  
  # ..............store key results...............
  brt_p_coords <- brt_res[["p_coords"]]
  brt_a_coords <- brt_res[["a_coords"]]
  brt_pred_stack <- brt_res[["brt_pred_stack"]]
  brt_gs_all <- brt_res[["brt_gs_all"]]
  brt_gs_reduced <- brt_res[["brt_gs_reduced"]]


  # ..............select best model...............
  brt_id_all <- which.max(brt_gs_all@results$test_AUC)
  brt_best_mod_all <- combineCV(brt_gs_all@models[[brt_id_all]])

  brt_id_reduced <- which.max(brt_gs_reduced@results$test_AUC)
  brt_best_mod_reduced <- combineCV(brt_gs_reduced@models[[brt_id_reduced]])

  # .............generate predictions.............
  brt_map_all <- predict(brt_best_mod_all, data = brt_pred_stack)
  brt_map_reduced <- predict(brt_best_mod_reduced, data = brt_pred_stack)

  rm(brt_res, brt_id_all, brt_best_mod_all, brt_id_reduced, brt_best_mod_reduced)


  # ...........vectorize coordinates .........
  brt_p_coords <- brt_p_coords %>% vect(geom = c("x", "y"), crs = "WSG84")
  brt_a_coords <- brt_a_coords %>% vect(geom = c("x", "y"), crs = "WSG84")


  # ........isolate all predictions.........
  pres_prediction_all <- extract(brt_map_all, brt_p_coords, xy = TRUE, ID = FALSE) %>%
    rename(prediction = lyr1) %>%
    mutate(pres_abs = 1,
           variables = "all")

  abs_prediction_all <- extract(brt_map_all, brt_a_coords, xy = TRUE, ID = FALSE) %>%
    rename(prediction = lyr1) %>%
    mutate(pres_abs = 0,
           variables = "all")


  # .......isolate reduced predictions......
  pres_prediction_reduced <- extract(brt_map_reduced, brt_p_coords, xy = TRUE, ID = FALSE) %>%
    rename(prediction = lyr1) %>%
    mutate(pres_abs = 1,
           variables = "reduced")

  abs_prediction_reduced <- extract(brt_map_reduced, brt_a_coords, xy = TRUE, ID = FALSE) %>%
    rename(prediction = lyr1) %>%
    mutate(pres_abs = 0,
           variables = "all")

  rm(brt_p_coords, brt_a_coords, brt_map_reduced, brt_map_all)

  # ............combine results..........
  combined_predictions <- rbind(pres_prediction_all, abs_prediction_all, pres_prediction_reduced, abs_prediction_reduced) %>%
    mutate(plot = i)

  brt_all_predictions <- rbind(brt_all_predictions, combined_predictions)

  rm(combined_predictions, pres_prediction_all, abs_prediction_all, pres_prediction_reduced, abs_prediction_reduced)

} # END BRT loop


```

## Select Variables - Iterate Through Plot 1-8

```{r}
# re-write data directory
rast_dir <- here("data", "raster-stacks", "full-stack")
point_dir <- here("data", "CampRoberts_spatial_data")

# create empty df
brt_all_predictions <- data.frame()


# run BRT for all 8 plots (all & reduced variables)
for (i in seq(1:2)) {
  
  # .............run both BRT models.............
  brt_res <- tune_brt(plot_number = i,
                      point_dir = point_dir,
                      rast_dir = rast_dir,
                      reduced_only = FALSE)
  
  # ..............store key results...............
  brt_p_coords <- brt_res[["p_coords"]]
  brt_a_coords <- brt_res[["a_coords"]]
  brt_pred_stack <- brt_res[["brt_pred_stack"]]
  brt_gs_all <- brt_res[["brt_gs_all"]]
  brt_gs_reduced <- brt_res[["brt_gs_reduced"]]
  
  
  # ..............select best model...............
  brt_id_all <- which.max(brt_gs_all@results$test_AUC)
  brt_best_mod_all <- combineCV(brt_gs_all@models[[brt_id_all]])
  
  brt_id_reduced <- which.max(brt_gs_reduced@results$test_AUC)
  brt_best_mod_reduced <- combineCV(brt_gs_reduced@models[[brt_id_reduced]])
  
  # .............generate predictions.............
  brt_map_all <- predict(brt_best_mod_all, data = brt_pred_stack)
  brt_map_reduced <- predict(brt_best_mod_reduced, data = brt_pred_stack)
  
  rm(brt_res, brt_id_all, brt_best_mod_all, brt_id_reduced, brt_best_mod_reduced)
  
  
  # ...........vectorize coordinates .........
  brt_p_coords <- brt_p_coords %>% vect(geom = c("x", "y"), crs = "WSG84")
  brt_a_coords <- brt_a_coords %>% vect(geom = c("x", "y"), crs = "WSG84")
  
  
  # ........isolate all predictions.........
  pres_prediction_all <- extract(brt_map_all, brt_p_coords, xy = TRUE, ID = FALSE) %>% 
    rename(prediction = lyr1) %>% 
    mutate(pres_abs = 1,
           variables = "all")
  
  abs_prediction_all <- extract(brt_map_all, brt_a_coords, xy = TRUE, ID = FALSE) %>% 
    rename(prediction = lyr1) %>% 
    mutate(pres_abs = 0,
           variables = "all")
  
  
  # .......isolate reduced predictions......
  pres_prediction_reduced <- extract(brt_map_reduced, brt_p_coords, xy = TRUE, ID = FALSE) %>% 
    rename(prediction = lyr1) %>% 
    mutate(pres_abs = 1,
           variables = "reduced")
  
  abs_prediction_reduced <- extract(brt_map_reduced, brt_a_coords, xy = TRUE, ID = FALSE) %>% 
    rename(prediction = lyr1) %>% 
    mutate(pres_abs = 0,
           variables = "all")
  
  rm(brt_p_coords, brt_a_coords, brt_map_reduced, brt_map_all)
  
  # ............combine results..........
  combined_predictions <- rbind(pres_prediction_all, abs_prediction_all, pres_prediction_reduced, abs_prediction_reduced) %>% 
    mutate(plot = i)
  
  brt_all_predictions <- rbind(brt_all_predictions, combined_predictions)
  
  rm(combined_predictions, pres_prediction_all, abs_prediction_all, pres_prediction_reduced, abs_prediction_reduced)
  
} # END for-loop

```

## Visualize Results
```{r}
test <- brt_all_predictions %>% 
  mutate(cor = ltm::biserial.cor(prediction, pres_abs))

ltm::biserial.cor(brt_all_predictions$prediction, brt_all_predictions$pres_abs)

cor(brt_all_predictions$prediction, brt_all_predictions$pres_abs, method = "pearson")
```

