---
title: "Final Run of All Plots"
author: "Briana Barajas"
date: 2024-05-08
---

## Preparation

```{r}
# wrangling
library(tidyverse)
library(here)
library(janitor)

# raster/geospatial
library(sf)
library(tmap)
library(terra)

# modeling
library(SDMtune)
library(rsample)
library(rJava)
```

```{r}
rm(list = ls())

# set data directory
rast_dir <- here("data", "raster-stacks")
point_dir <- here("data", "CampRoberts_spatial_data")

# source functions
source(here("R", "sdmtune-maxent-final-function.R"))

# occurrence points
points <- read_sf(here(point_dir, "Species_pts", "CR_BASP_obs_11Jul22.shp"), quiet = TRUE) %>% 
  clean_names() %>% 
  filter(plot == 1)
```


## Plot Summary

Each plot indicates a different transect where cover board surveys were conducted, and contains it's own unique values of occurrence points. The maximum entropy model is often noted for it's capabilities even when given a small data set. Since each plot has a unique number of occurrence points, some of the analyses conducted compare plots with low vs. high numbers of points to see how the AUC was affected.

| Plot Number | Number of Occurrences |
|-------------|----------------------|
| 1           | 51                   |
| 2           | 57                   |
| 3           | 78\*                 |
| 4           | 19                   |
| 5           | 72                   |
| 6           | 39                   |
| 7           | 10\*                 |
| 8           | 21                   |

## Test MaxEnt Function
```{r}
# run model
maxent_res <- tune_maxent(plot_number = 1,
                          point_dir = point_dir,
                          rast_dir = rast_dir,
                          k_folds = 3)

# pull models from results
maxent_test_obj <- maxent_res[["maxent_test"]]
maxent_pred_stack <- maxent_res[["maxent_pred_stack"]]
maxent_gs <- maxent_res[["maxent_gs"]]
maxent_reduced_model <- maxent_res[["maxent_mod_reduced"]]
maxent_bg_coords <- maxent_res[["bg_coords"]]

# isolate the top performing model & predict
maxent_id <- which.max(maxent_gs@results$test_AUC)
maxent_best_mod <- combineCV(maxent_gs@models[[maxent_id]])
maxent_map <- predict(maxent_best_mod, data = maxent_pred_stack,
                      type = "cloglog",
                      clamp = TRUE)

# plot results
plotROC(maxent_best_mod, test = maxent_test_obj)

tm_shape(maxent_map) +
  tm_raster(midpoint = 0) +
  tm_shape(points) +
  tm_dots(size = 0.1) +
  tm_layout(legend.position = c(1, 0.2))

```

## Extract Values

```{r}
# extract absence values (bg coords)
abs_points <- maxent_bg_coords %>% 
  st_as_sf(coords = c("x", "y"), crs = "WSG84")
  # vect(geom = c("x", "y"), crs = "WSG84")

# extract values for presence points
pres_points <- points %>% 
  select(latitude, longitude)

tm_shape(abs_points) +
  tm_dots(col = "lightblue", size = 0.1) +
  tm_shape(pres_points) +
  tm_dots(col = "navy", shape = 17, size = 0.15)
```


```{r, eval=FALSE}
### FOR BRT ### 

# extract absence prediction values
abs_points <- pres_abs_points %>% 
  filter(basp_pa == 0) %>% 
  select(longitude, latitude) %>% 
  vect(geom=c("longitude", "latitude"), crs="WSG84")

abs_predicted <- extract(brt_map, abs_points, xy = TRUE, ID = FALSE) %>% 
  mutate(basp_pa = 0) %>% 
  rename(prediction = lyr1)

# extract presence prediction values
pres_points <- pres_abs_points %>% 
  filter(basp_pa == 1) %>% 
  select(longitude, latitude) %>% 
  vect(geom=c("longitude", "latitude"), crs="WSG84")

pres_predicted <- extract(brt_map, pres_points, xy = TRUE, ID = FALSE) %>% 
  mutate(basp_pa = 1) %>% 
  rename(prediction = lyr1)


# combine
abs_predicted %>% 
  rbind(pres_predicted) %>% 
  mutate(model = "BRT reduced")
```


